### 2017-02-23 thesis/latex/thesis.tex

\subsection{How does dynein walk?}

Cianfrocco \textit{et. al} \cite{cianfroccoreview} propose dynein achieves motion via transitioning through the cycle of states shown in Fig. \ref{mech-cycle}. This model will hereon be referred to as the Cianfrocco model. The model claims the primary events in the dynein step are as follows. First, starting with both binding domains attached to the MT, ATP binds to the motor domain, which causes it to change conformation. ATP hydrolysis to ADP-Pi causes further conformation changes. These changes cause the binding domain to release the microtubule, meaning only one binding domain is MT-bound. The free binding domain then undergoes another conformation change known as the ``priming stroke,'' which changes the orientation of the binding domain with respect to the MT. Finally the binding domain rebinds to the MT, causing a final conformation change in the motor which returns it to the beginning of the cycle. The exact order of these events is not precisely known. More detail on each step of the cycle is given below.\\

\textbf{Things which are known about dynein which support the Cianfrocco model}
\begin{itemize}
\item AAA1 only ATPase necessary for motility, which is why we don't include the other ATPases in our model \cite{cianfroccoreview}
\item AAA5 mediates stalk register changes Schmidt H, Zalyte R, Urnavicius L, Carter AP. 2014. Structure of human cytoplasmic dynein-2 primed for
  its power stroke. Nature 518:435–38
\item Linker moves as response to AAA1 nucleotide binding Burgess SA, Walker ML, Sakakibara H, Knight PJ, Oiwa K. 2003. Dynein structure and power stroke. Nature
  421:715–8
\item Crystal structures with Vanadium-Adp show dynein in the pre-powerstroke state, hinting that ADP-Pi-bound dynein takes on a different conformation than ADP-Pi free dynein
\end{itemize}

\subsubsection{ATP binding / unbinding effects}
\subsubsection{Microtubule un/rebinding}
Microtubule binding and unbinding play an important role in the dynein mechanochemical cycle. The MTBD/stalk system has two conformational states: a high-MT-affinity state where MT binding occurs at a fast rate, and a low-affinity state with a slower binding rate \. The MTBD/stalk alternates between these two states as the motor progresses through the mechanochemical cycle

\begin{figure}[h]
  \centering
  \includegraphics[width=.65\textwidth,keepaspectratio]{../../figures/mtbd-complex.png}
  \caption{Crystal structure of the MTBD-MT complex. Modified from Redwine \textit{et. al} \cite{redwineMTBDcomplex}. Is this figure actually useful, or would a diagram be better?}
  \label{dynein-artist-rendition-2}
\end{figure}

\subsubsection{Domain movement}
Dynein domains have been shown to move in an alternating fashion, with motor domains moving twice the distance of head domains. This indicates an alternate shuffling-type movement.\\

Flesh this out? Make a table of lots of different experiments and what they indicate?\\

\subsubsection{Conformation change of motor (``priming stroke'')}
A FRET (fluorescent resonance energy transfer) study, which probes the closeness of different labeled domains on a protein, showed that dynein is found in two different major conformations, depending on the nucleotide bound to its AAA1 ATPase site \cite{FRETstatepaper}. When no nucleotide, ATP or ADP is bound, the authors label the protein ``state I.'' When ADP-Pi bound, the motor is said to be in ``state II,'' as shown in Fig \ref{nucleotide-state}. In addition, motors in the presence of the unhydrolyzable ATP analog were found in state II, indicating the transition from state I to II occurs during ATP binding. Similar reasoning shows that the state II to I transition occurs when the protein is ADP bound. This allows the authors to create the state diagram shown in Fig \ref{nucleotide-state-diagram}.\\

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../../figures/two-state-em-diagram}
%%   \caption{}
%%   \label{nuecleotide-state}
%% \end{figure}

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../../figures/nucleotide-state-cycle}
%%   \caption{}
%%   \label{nucleotide-state-diagram}
%% \end{figure}

This single FRET study illustrates how the conformational aspect of the mechanochemical cycle works. Dynein begins in state I and binds ATP into its AAA1 ATPase, changing into state II by changing its stalk-tail angle to move the tail outward. ATP is then hydrolized into ADP-Pi. Pi is then ejected from the AAA1 site, leaving ADP behind. This causes the reverse state transition, moving the tail inward. When combined with microtubule affinity changes, this primitive model of the dynein cycle could explain how dynein walks.\\

\textit{It is interesting to note that the hydrolysis of ATP is not actually used to directly power the motor -- binding AMP-PMP, a nonhydrolyzable analog of ATP, can also cause the motor to change from state I to state II \cite{FRETstatepaper}. The specialness of ATP is that it can be broken down into ADP, which causes another conformation change back to state I. Thus, something something the transfer of energy from nucleotide to motor is in its shape, not in its chemistry.}\\

\textit{Add figure somewhere of this cycle superimposed with markings of what states are considered onebound and bothbound}\\

\textbf{more on how ATP actually causes the conformation change, if that's available. The change from an open/closed ATPase is what changes the nucleotide affinity, go into depth on that, but actual info on how ATP binding causes conformation change would be good. Something about going from a planar to nonplanar conformation?}\\

also how crystal structures are in different positions

\subsubsection{Experimental results on dynein’s step}
Weihong's figures go here.\\

\subsubsection{c What is still needed to support theory}
Where our simulation fits in. Why we need this kind of information.
There are many different things which might be important for dynein's walk which we don't take into account: springiness of tail/stalk, changes in spring constants, repulsion between the heads, multiple pre/post powerstroke conformations. We are trying to see if these things are actually important.

\subsubsection{Other theories?}
Imamula model: how it is different from Cianfrocco model, and why we chose the Cianfrocco model.\\
	\subsection{Past Simulations}
		\subsubsection{Sarlah Winch Model}
		\subsubsection{Other models}

\subsection{Moving forward}
Talk about why simulation is the only way to verify the mechanochemical cycle, since experiments
aren't high resolution enough to make good videos of the walk, and even if they were, in some sense
knowing that dynein behaves identically to a simple mathematical model is more informative than
a video, since there is lots of information in a video and really precise info in a simulation.\\

\textbf{Not sure where this goes:} The 'goal' of our experiment, at a very high level, is to
understand how dynein works. We can know this if we can come up with the minimum set of characteristics a physical system needs to act like dynein. This information is the same thing as knowing
``how dynein walks.'' We don't need to know all the complex dynamics of every single amino acid in
dynein, we only need to know our rough model.\\

\section{Chapter: Introduction II}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.65\textwidth,keepaspectratio]{../../figures/literature/burgess-dynein-angles}
  \caption{Cryo-electron microscopy images of dynein in two different nucleotide states. Left: dynein frozen in ADP-Pi state. Middle: dynein without nucleotide. Right: two states superimposed on one another. Bottom: cartoons representing how dynein may bind to the MT in either state. From \cite{burgess-paper}.}
  \label{fig:cryo-em-angles}
\end{figure}

@@

\subsubsection{Stepping pattern}
Several studies have been done on the dynein motor's stepping pattern using selectively-labeled motor domains \cite{reck2006single} \cite{weihongpaper}. These studies suggest that dynein has a

>>>>

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.65\textwidth,keepaspectratio]{../../figures//burgess-dynein-angles}
  \caption{Cryo-electron microscopy images of dynein in two different nucleotide states. Left: dynein frozen in ADP-Pi state. Middle: dynein without nucleotide. Right: two states superimposed on one another. Bottom: cartoons representing how dynein may bind to the MT in either state. From \cite{burgess-paper}.}
  \label{fig:cryo-em-angles}
\end{figure}

@@

\subsubsection{Stepping pattern}
Several studies have been done on the dynein motor's stepping pattern using selectively-labeled motor domains \cite{reck2006single} \cite{weihongpaper}. These studies show dynein's stepping pattern to be stochastic and weakly forward-directed. As shown in Figure (\ref{fig:weihong-steps}), dynein takes many backwards steps, and also many steps with components perpendicular to the microtubule. This indicates the process dynein uses to generate motion has a stochastic component to it. It is likely that the physical states of subcomponents in the motor, and/or the effects of Brownian motion due to water particle collision, play a role in dynein's stepping mechanism.\\

\begin{figure}[h!]
  \centering
  \begin{minipage}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{../../figures/weihong-step-size}
    \caption{Size of steps taken by dynein motor parallel to microtubule. Positive axis corresponds to steps towards the microtubule minus-end. Figure modified from \cite{weihongpaper}.}
  \end{minipage}
  \hspace{.2cm}
  \begin{minipage}[b]{0.45\textwidth}
    \includegraphics[width=\textwidth]{../../figures/weihong-off-axis-steps}
    \caption{Size of ``off-axis'' steps taken by the dynein motor perpendicular to the microtubule. Figure modified from \cite{weihongpaper}.}
  \end{minipage}
  \label{fig:weihong-steps}
\end{figure}

\subsection{Dynein models}
Data presented in the previous sections paints a

\subsubsection{Cianfrocco Model}
Cianfrocco \textit{et. al} \cite{cianfroccoreview} propose dynein achieves motion via transitioning through the cycle of states shown in Fig. \ref{mech-cycle}. This model will hereon be referred to as the Cianfrocco model. The model claims the primary events in the dynein step are as follows. First, starting with both binding domains attached to the MT, ATP binds to the motor domain, which causes it to change conformation. ATP hydrolysis to ADP-Pi causes further conformation changes. These changes cause the binding domain to release the microtubule, meaning only one binding domain is MT-bound. The free binding domain then undergoes another conformation change known as the ``priming stroke,'' which changes the orientation of the binding domain with respect to the MT. Finally the binding domain rebinds to the MT, causing a final conformation change in the motor which returns it to the beginning of the cycle. The exact order of these events is not precisely known. More detail on each step of the cycle is given below.\\

\subsubsection{Other models}
If there are other models which conflict with Cianfrocco, here would be a place to put them.\\

\subsection{This work}
\subsubsection{Our model}
what we're trying to do with our project

\textit{Add a figure (maybe here?) of all the events which are known to happen in dynein, along with our mechanochemical cycle. It would basically show all events occuring at two times: unbinding or rebinding.}\\

\subsubsection{Goals}
What we hope to gain by simulating such a model, and what it would contribute to the current literature

%% \textbf{Things which are known about dynein which support the Cianfrocco model}
%% \begin{itemize}
%% \item AAA1 only ATPase necessary for motility, which is why we don't include the other ATPases in our model \cite{cianfroccoreview}
%% \item AAA5 mediates stalk register changes Schmidt H, Zalyte R, Urnavicius L, Carter AP. 2014. Structure of human cytoplasmic dynein-2 primed for
%%   its power stroke. Nature 518:435–38
%% \item Linker moves as response to AAA1 nucleotide binding Burgess SA, Walker ML, Sakakibara H, Knight PJ, Oiwa K. 2003. Dynein structure and power stroke. Nature
%%   421:715–8
%% \item Crystal structures with Vanadium-Adp show dynein in the pre-powerstroke state, hinting that ADP-Pi-bound dynein takes on a different conformation than ADP-Pi free dynein
%% \end{itemize}

%% \textit{It is interesting to note that the hydrolysis of ATP is not actually used to directly power the motor -- binding AMP-PMP, a nonhydrolyzable analog of ATP, can also cause the motor to change from state I to state II \cite{FRETstatepaper}. The specialness of ATP is that it can be broken down into ADP, which causes another conformation change back to state I. Thus, something something the transfer of energy from nucleotide to motor is in its shape, not in its chemistry.}\\
